---
title: "10x Genomics"
output: html_notebook
---


```{r}
library(dplyr)
library(anndata)
library(reticulate)
Sys.setenv(RETICULATE_PYTHON = #path to python )
library(ggplot2)
library(tools)
library(stats)
library(sjmisc)
library(stringr)
library(parallel)
library(doParallel)
library(foreach)
```
```{r}
detectCores(logical = FALSE)
ls("package:parallel")
ls("package:doParallel")
ls("package:foreach")
```


# Determine how many genes are expressed in each cell

analysis <- function(filename) {
  adata <- anndata::read_loom(filename)
  adata
  gpc <- data.frame(CellID = adata$obs, num_genes = as.integer(NA), library_size = as.integer(NA))     # create a new dataframe to hold number of genes in each cell
  for (i in 1:length(adata$X[,1])) {
    genes<-adata$X[i,]
    gpc[i, "num_genes"] <- length(genes[genes > 0])
    gpc[i, "library_size"] <- sum(genes)
  }
  temp <- c(file_path_sans_ext(filename), "csv")
  savfile <- paste(temp, collapse = ".")
  write.table(gpc, file=savfile)
}



# Plotting results for number of read genes per cell

denplot <- function(gdata, numgenes, libsize, filename) {
  temp <- c(file_path_sans_ext(filename), "_reads", ".png")
  greads <- paste(temp, collapse = "")
  png(filename = greads, width=600, height=350, type = "cairo")
  hist(numgenes,col = "orange", border = "black", probability = TRUE,
     xlab = "num_genes", main = "Distribution of Genes Expressed Per Cell", xlim = c(0,max(numgenes)+2000), breaks = 100)
  lines(density(numgenes), lwd = 3, col = "black")
  dev.off()
  
  temp <- c(file_path_sans_ext(filename), "_libsize", ".png")
  glibsize <- paste(temp, collapse = "")
  png(filename = glibsize, width=600, height=350, type = "cairo")
  hist(libsize,col = "orange", border = "black", probability = TRUE,
     xlab = "library size", main = "Distribution of Library Size Per Cell", xlim = c(0,max(libsize)+10000), breaks = 116)
  lines(density(libsize), lwd = 3, col = "black")
  dev.off()
  
}



# Finding Pearson Correlation Coefficient and plotting a scatter plot two show relationship between number of genes and library size

scatterpcc <- function(dataframe, numgenes, libsize, filename) {
  temp <- c(file_path_sans_ext(filename), "_pcc_scat", ".png")
  filename <- paste(temp, collapse = "")
  png(filename = filename, width=600, height=350, type = "cairo")
  r <- cor(numgenes, libsize, use = "everything", method = "pearson")   # Find PCC
  r <- format(round(r, 4), nsmall = 4)    # Format PCC to 4 decimal places
  plot(numgenes, libsize, 
     main = "Number of Genes vs. Library Size", xlab = "Number of Genes",
     ylab = "Library Size", pch = 19, frame.plot = TRUE, col = "orange",)   # Scatter Plot
  temp <- c("r = ", r)
  pcc <- paste(temp, collapse = "")
  legend("topleft", c("Pearson Correlation", "Coefficient", pcc))     # Add Legend to Plot
  abline(lm(libsize ~ numgenes, data = dataframe), col = "black")   # Add Linear Regression to scatter plot
  dev.off()
}

```{r}
# Finding correlations between each TF and TG for each loom file, and saving results to matrix

tftg_corr <- function(set_list ,genes, trans_facs, settings, index) {
      # Matrix of Transcription Factors (TF/s) as variables and Targeted Genes (TG/s) as observations
  pcorr <- matrix(0, nrow = nrow(genes), ncol = trans_facs$n_vars)
  colnames(pcorr) <- trans_facs$var_names
  rownames(pcorr) <- genes$V4
  head(pcorr)

# --------------------------------------------------------------------------------------
        # This code is slow, it takes for 4-5 hours to move to the next iterator for loop B because loop C has around 50,000 iterations and loop B is around 250 iterations
  
  for (a in 1:length(set_list)) {
    loom <- read_loom(set_list[[a]])
    print("we read the first file")
    maxx <- trans_facs$n_vars
    for(b in 1:maxx) {                                  # LOOP B
      print("we are looping through the TFs")
      TF_name <- trans_facs$var_names[b]                # saving TF name
      for (i in 1:loom$n_vars) {                        # LOOP C
        print("we are looping through the TGs")
        k <- loom$var_names[i]                          # saving TG  name
        if (k %in% genes$V4 == TRUE && k != TF_name) {            # Verifying if TG exists in bed file and TG != TF
          gexp <- loom$X[,k]
          TF_exp <- loom$X[,TF_name]
          if (length(gexp[gexp>0]) > 3) {             # Pearson Correlation (PC) is only calculated if TG is read in more than 3 cells
            print("performing correlation")
            r <- cor(TF_exp, gexp, use = "everything", method = "pearson")
            print(r)
          }
          else {
            r <- "NULL"               # If TG is not expressed in at least 3 cells          
            print(r)
          }
        }
        else {
          r <- "NULL"              # If they are the same gene 
          print(r)
        }
        pcorr[k, TF_name] <- r
        print("getting next gene")
      }
      print("getting next TF")
    }
    print("getting next loom")
  }
  colnames(corr) <- trans_facs$var_names
  rownames(corr) <- genes$V4
  if (length(settings) != 0) {
    temp <- c("correlations", settings[a])
    temp <- paste(temp, collapse = "_")
    temp <- c(temp, ".csv")
    savfile <- paste(temp, collapse = "")
  }
  write.csv(pcorr, file = savfile)
}
```




files <- list.files(#path to file , pattern = "*.loom")
files
for (i in 1:length(files)) {
  bool <- str_contains(files[i], "vsmc", ignore.case = TRUE)
  if (bool == TRUE) {
    analysis(files[i])
  }
}
```


files <- list.files(#path to file, pattern = "*.csv")
for (i in 1:length(files)) {
  bool <- str_contains(files[i], "vsmc", ignore.case = TRUE)
  if (bool == TRUE) {
    gpc <- read.table(files[i])
    denplot(gpc, gpc$num_genes, gpc$library_size, files[i])
    scatterpcc(gpc, gpc$num_genes, gpc$library_size, files[i])
  }
}
```

```{r}
files <- list.files(#path to file , pattern = "*.loom")
genes_bed <- read.table(#path to file , header = FALSE)
TFs <- read_csv(#path to file)

# Settings to be analyzed saved in a list must be insert as 
    #genes type (Wild or KO), reading week, HFD (if exist, if not then do not include), all separated by _
sets <- list("wt_3w", "ko_3w", "wt_6w", "ko_6w", "wt_1w_HFD", "wt_12w_HFD", "ko_12w_HFD")

sp_sets <- list()
k = 1
for ( i in 1:length(sets)) {
  for ( j in 1:length(files)) {
    bool <- str_contains(files[j], sets[i], ignore.case = TRUE)
    if (bool==TRUE) {
      vsmc <- str_contains(files[j], "vsmc", ignore.case = TRUE)
      if (vsmc == FALSE) {
        sp_sets[k] <- files[j]
        if (any(sp_sets == "NULL")) {
          sp_sets <- sp_sets[-which(sapply(sp_sets, is.null))]                     # removes null values from list
        }
      }
    }
    k = k+1
  }
}
print("--------------------")
sp_sets
```

```{r}
# Scratch work to enhance speed of code (Not Complete)

cluster <- parallel::makeCluster((detectCores()-20))
  doParallel::registerDoParallel(cluster)

  loom <- read_loom(sp_sets[[1]])
  print("we read the first file")
  maxx <- TFs$n_vars
  temp <- list()
    foreach(j=100:maxx, .combine = 'cbind') %dopar% {
      TF_name <- TFs$var_names[j]                # saving TF name
      for (i in 1:50) {
        k <- loom$var_names[i]                   # saving TG  name && k != TF_name
        #k <- genes_bed$V4[i]
        #if (k %in% genes_bed$V4 == TRUE) {            # Verifying if TG exists in bed file and TG != TF
        gexp <- loom$X[,k]
        TF_exp <- loom$X[,TF_name]
        #temp[i] <- 3 * i
        gexp <- gexp[ ,gexp[,1]>0]            # Pearson Correlation (PC) is only calculated if TG is read in more than 3 cells
        #r <- cor(TF_exp, gexp, use = "everything", method = "pearson")
        temp[i] <- gexp[5]
      }
          #else {
            #r <- 2               # If TG is not expressed in at least 3 cells
            #test[i] <- 3
          #}
      temp
    }
        #else {
          #r <- 2              # If they are the same gene
          #test[i] <- 3
        #}
      #}
    #j <- test
```
